Toward nonlinear multigrid \\ for nonlinear and nonlocal variational inequalities

### 2023-02-09 talk/slides.tex
a bit of progress ... now for a practice run

\bigskip
\begin{itemize}
\item what is needed for scalable multilevel solutions?
    \begin{enumerate}
    \item<2>[1.] iterates must be admissible
    \item<2>[2.] global linearization must be avoided
    \item<2>[3.] smoother cost must be comparable to one residual    
    \end{enumerate}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item semi-smooth Newton can also yield mesh-independent iterations (Farrell et al.~2020)
\item other VI multilevel strategies:
    \begin{itemize}
    \item[$\circ$] projected FAS multigrid for linear CPs (Brandt \& Cryer, 1983)
    \item[$\circ$] monotone multigrid (Kornhuber, 1994)
    \item[$\circ$] multilevel constraint decomposition (Tai, 2003)
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}{multigrid strategies for nonlinear VIs: feature table}

{\small
\begin{tabular}{l|ccccc}
   & \stacktwo{admissible}{iterates} & \stacktwo{mesh-indep.}{rates}
       & \stacktwo{no global}{linearization}
         & \stacktwo{PETSc or}{Firedrake}  \\ \hline
RS NM                      & $\checkmark$ &  & & $\checkmark$ \\
\quad $+$ NI & $\checkmark$ & $\checkmark$ & & $\checkmark$ \\
SS NM             &              & $\checkmark$ & & $\checkmark$ \\
{\color{FireBrick} FASCD}  & $\checkmark$ & {\color{FireBrick} \textbf{?}} & $\checkmark$ & {\color{FireBrick} {\large $\checkmark$}}
\end{tabular}
}

@@

CD gloss

@@

\begin{itemize}
\item x
\end{itemize}

\includegraphics[width=0.5\textwidth]{images/icedecomp.png}
\end{frame}

@@

pseudocode
\end{frame}

@@

\begin{itemize}
\item 1D SIA
\end{itemize}

@@

\begin{itemize}
\item x
\end{itemize}

\includegraphics[width=0.5\textwidth]{images/fascd-firedrake-dome.png}
\end{frame}

>>>>

\begin{itemize}
\item[]
    \begin{itemize}
    \item[$\circ$] for more on this problem see (Bueler, 2021)
    \end{itemize}
\end{itemize}

\begin{itemize}
\item {\color{FireBrick} what is needed for scalable multilevel solutions?}
    \begin{enumerate}
    \item<2>[1.] iterates must be admissible
    \item<2>[2.] global linearization of $\Phi(s)$ must be avoided
    \item<2>[3.] smoother cost must be comparable to one residual    
    \end{enumerate}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item semi-smooth Newton also yields mesh-independent iterations
    \begin{itemize}
    \item[$\circ$] penalty scaling argument (Farrell et al.~2020)
    \end{itemize}
\item other VI multilevel strategies:
    \begin{itemize}
    \item[$\circ$] projected FAS multigrid for linear CPs (Brandt \& Cryer, 1983)
    \item[$\circ$] monotone multigrid (Kornhuber, 1994)
    \item[$\circ$] multilevel constraint decomposition (Tai, 2003)  \hfill $\gets$ \emph{more below}
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}{multigrid strategies for VIs: feature table}

{\small
\begin{tabular}{l|ccccc}
   & \stacktwo{admissible}{iterates} & \stacktwo{mesh-indep.}{rates}
       & \stacktwo{no global}{linearization}
         & \stacktwo{PETSc or}{Firedrake}  \\ \hline
RS NM                      & $\checkmark$ &  & & $\checkmark$ \\
\quad $+$ NI & $\checkmark$ & $\checkmark$ & & $\checkmark$ \\
SS NM             &              & $\checkmark$ & & $\checkmark$ \\
{\color{FireBrick} FASCD}  & $\checkmark$ & {\color{FireBrick} \textbf{?}} & $\checkmark$ & \only<2>{{\Large {\color{FireBrick} $\checkmark$}}} {\Large \strut}
\end{tabular}
}

@@

\begin{itemize}
\item Tai's (2003) \emph{constraint decomposition} (CD) for VIs follows subspace decomposition for PDEs (Xu 1992)
\item 
\end{itemize}
\end{frame}

\begin{frame}{as decomposition of the fluid layer}

@@

\begin{itemize}
\item what does 
\end{itemize}

\begin{center}
\includegraphics[width=\textwidth]{images/icedecomp.png}
\end{center}
\end{frame}

@@

% case of lower obstacle only
\begin{pseudo}
\pr{fascd-vcycle}(\ell^J,\underline{\gamma}^J; w^J)\text{:} \\+
    $\underline{\chi}^J = \underline{\gamma}^J - w^J {\large \strut}$ \\
    %$\overline{\chi}^J = \overline{\gamma}^J - w^J {\large \strut}$ \\
    for $j=J$ downto $j=1$ \\+
      $\underline{\chi}^{j-1} = \maxR \underline{\chi}^j {\large \strut}$ \\
      %$\overline{\chi}^{j-1} = \minR \overline{\chi}^j {\large \strut}$ \\
      $\underline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1} {\large \strut}$ \\
      %$\overline{\phi}^j = \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \\
      $y^j = 0$ \\
      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,w^j;y^j)$ \hspace{8mm} \ct{smoothing in $\mathcal{D}^j$}\\
      $w^{j-1} = \iR(w^j + y^j)$ \\
      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \\-
    $z^0 = 0$ \\
    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,w^0;z^0)$ \hspace{10mm} \ct{coarse solve in $\mathcal{U}^0$} \\
    for $j=1$ to $j=J$ \\+
      $z^j = y^{j} + P z^{j-1}$ \\
      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,w^j;z^j)$ \hspace{8mm} \ct{smoothing in $\mathcal{U}^j$} \\-
    $w^J \gets w^J+z^J$
\end{pseudo}
\end{frame}

\begin{itemize}
\item preliminary results
\item dome test case in lubrication approximation
    \begin{itemize}
    \item[$\circ$] \emph{shallow ice approximation} = SIA
    \item[$\circ$] $\Phi(s) = -\bu|_s \cdot \bn_s$ is a differential operator
    \item[$\circ$] $s^{8/3} \in W^{1,p}(\Omega)$ but not in $C^2$
    \end{itemize}
\item FASCD algorithm using a strong smoother (\texttt{vinewtonrsls})
    \begin{itemize}
    \item[$\circ$] Firedrake $P_1$ elements
    \end{itemize}
\end{itemize}

\begin{center}
\includegraphics[width=0.7\textwidth]{images/fascd-firedrake-dome.png}
\end{center}
\end{frame}

@@

\begin{itemize}
\item again in the lubrication approximation,
\end{itemize}

@@
